A First Effort for a Distributed Segment-based Approach on Self-Assembled Nano Networks

### 2013-10-16 trunk/PAPER/introduction.tex
Adding missing reference

In the last years, Self-assembled nanoscale architectures~\cite{winfree1998, yan2003}
are emerging as a promising technology due their tera/peta scale of
integration, defect tolerance tolerance and huge potential computing
capabilities. These technologies are certainly still at
their early stage of development, however different laboratory demos and
prof-of-concepts architectures have been presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1, pistol2009}.
The main idea behind this approach is to exploit the physical regularity and
stabiliy of DNA structures in order to create a scaffold onto which
nano-devices (e.g. nanowires and CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be
attached. This can be achieved by designing approapriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see figure~\ref{fig:dna_tag}.

\begin{figure}
  \centering
    \includegraphics[width=0.30\textwidth]{pictures/dna2b.eps}
  \caption{DNA tag matching}
  \label{fig:dna_tag}
\end{figure}

A detailed description of the chemical properties involved is far
beyond the scope of this paper (see also~\cite{braun1998, seeman1999}), so we will focus on
the three main challenges that this new fabrication process introduces
in Computer Design: (i)limited node complexity, (ii) large scale
randomness and (iii) high defect rates.

@@

Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top down design , the same nature of a bottom up self
assembled process cannot assume such a deterministic device placement
process.

These aspects of a DNA-self assembly process lead to some important
implications to be addressed when approaching to the design of
computer architecture: a computational model should be founded on a
distributed network of small computing and storage nodes, randomly
placed and interconnected (see figure~\ref{fig:nana}).
Proof-of-concept of such kind of computational model can be found for
example in~\cite{patwardhan2006_1}.  Since no regularity can be
assumed in such network, a topology agnostic strategy that avoids
deadlock should be adopted in order to route packets (e.g.
instructions and data operands) among nodes.  It is important to
underline that, altought we can assume an architecture similar to the
one depicted in~\ref{fig:nana} as background, the topology discovery
problem is more general and not stricly dependent on any of the
underlying computing architectures.

@@

The main contribution presented in this paper is to bring the advantages of
a segment-based approach to deadlock freedom without requiring any topology graph,
external defect map or centralized algorithm execution.  We introduce DiSR, a first effort for a Distributed
Segment-based Routing in large scale DNA self assembled networks.
While maintaining similar properties to segment-based , the DiSR
approach is not intended to discover the "optimal" segment choice
(ideally reachable with the knowledge of the topology graph) but just
to demonstrate a concrete model that can fit into such complex,
irregular and large sized topologies.
Given the irregular/high-defect rates scenario discussed above, is essential
avoiding non-scalable or resource hungry solutions such as
centralized off-chip algorithms, tree-based approaches, virtual channels
or expensive implementations based on hardware redundancy.

The paper is organized as follows: in Section~\ref{sec:related_works}
we summarize the main approaches to topology agnostic routing. Next in
Section~\ref{sec:disr_concepts} are described the main elements of the
proposed approach, showing a the DiSR execution model in
Section~\ref{sec:execution_model}.  In Section~\ref{sec:simulation}
simulations are carried out with the opensource tool Nanoxim to
demostrate how DiSR preserves some of the main segment-based
properties and how compares to the tree-based approaches. Finally a
draft of a architectural implementation is shown in
Section~\ref{sec:implementation} to evaluate the feasibility of the
approach in terms of node complexity.

\section{Related Works}
\label{sec:related_works}
Several works~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003} have been presented addressing the problem of topology
agnostic routing. While they show interesting
performances, the requirement for virtual channels resources 
to avoid deadlock makes them not suitable for the specific DNA self
assembled networks we are assuming as background.
Other topology independent algorithms~\cite{schroeder1991, koibuchi2001, cherkasova1996} do not use virtual channels to
achieve deadlock freedom, but turn proibitions. 
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, then placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
While the hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
striclty depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{dally1994, duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, sui2000, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010, kim2006, park2006}.

In~\cite{mejia_ipdps06} authors present SR, an approach the solves these
limitations by allowing turn restrictions to be placed locally,
independently from other restrictions. The whole network is
partitioned into segments, and a each bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom and connected networks. This \emph{locality
independence} property, together with no requirement of any particular tree/root
choice, would make it the best choice for the given scenario;
however, its topology independency still requires the knowledge of the
whole network graph in order to find the segments.

The DiSR approach presented in this work is a first attempt of
reaching the same goal of a segment based approach, that is
estabilishing segments, without requiring any topology graph or
centralized execution.

>>>>

In the last years, Self-assembled nanoscale architectures~\cite{winfree1998, yan2003}
are emerging as a promising technology due their tera/peta scale of
integration, defect tolerance and huge potential computing
capabilities. These technologies are certainly still at
their early stage of development, however different laboratory demos and
prof-of-concepts architectures have been presented~\cite{patwardhan2004, patwardhan2006_1, pistol2009}.
%, patwardhan2006
The main idea behind this approach is to exploit the physical regularity and
stability of DNA structures in order to create a scaffold onto which
nano-devices (e.g. nanowires and CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be
attached. This can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold.
%(see figure~\ref{fig:dna_tag}.
%
%\begin{figure}
%  \centering
%    \includegraphics[width=0.30\textwidth]{pictures/dna2b.eps}
%  \caption{DNA tag matching}
%  \label{fig:dna_tag}
%\end{figure}
A detailed description of the chemical properties involved is far
beyond the scope of this paper (see also~\cite{braun1998, seeman1999}), so we will focus on
the three main challenges that this new fabrication process introduces
in Computer Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.

@@

Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top down design , the same nature of a bottom up self
assembled process cannot assume such a deterministic device placement
process, thus defect tolerance is something like a design requirement
more than an exception to be avoided.

These aspects of a DNA-self assembly process lead to some important
implications to be addressed when approaching to the design of
computer architecture: a computational model should be founded on a
distributed network of small computing and storage nodes, randomly
placed and interconnected (see figure~\ref{fig:nana}).
Proof-of-concept of such kind of computational model can be found for
example in~\cite{patwardhan2006_1}.  Since no regularity can be
assumed in such network, a topology agnostic strategy that avoids
deadlock should be adopted in order to route packets (e.g.
instructions and data operands) among nodes.  It is important to
underline that, altougth we can assume an architecture similar to the
one depicted in Figure~\ref{fig:nana} as background, the topology discovery
problem faced in this work is more general and not strictly dependent on any of the
underlying computing architecture.

@@

In this paper we introduce DiSR, a first effort for a Distributed
Segment-based Routing in large scale DNA self assembled networks.
The main contribution is to bring the advantages of
a segment-based approach~\cite{mejia_ipdps06} to deadlock freedom without requiring any topology graph,
external defect map or centralized algorithm execution. 
While maintaining similar properties to segment-based , the DiSR
approach is not intended to discover the "optimal" segment choice
(ideally reachable with the knowledge of the topology graph) but just
to demonstrate a concrete model that can fit into such complex,
irregular and large sized topologies.
Given the irregular/high-defect rates scenario discussed above, is essential
avoiding non-scalable or resource hungry solutions such as
centralized off-chip algorithms, tree-based approaches, virtual channels
or expensive implementations based on hardware redundancy.

The paper is organized as follows: in Section~\ref{sec:related_works}
we summarize the main approaches to topology agnostic routing. Next in
Section~\ref{sec:disr_concepts} are described the main elements and
the execution model of the proposed approach.  In Section~\ref{sec:simulation}
simulations are carried out with the opensource tool Nanoxim to
demonstrate how DiSR preserves some of the main segment-based
properties and how compares to the tree-based approaches. Finally a
draft of a architectural implementation is shown in
Section~\ref{sec:implementation} to evaluate the feasibility of the
approach in terms of node complexity.

\section{Related Works}
\label{sec:related_works}
Several works~\cite{sancho2002, skeie2002, skeie2004, koibuchi2003} have been presented addressing the problem of topology
agnostic routing. While they show interesting
performances, the requirement for virtual channels resources 
to avoid deadlock makes them not suitable for the specific DNA self
assembled networks we are assuming as background.
Other topology independent algorithms~\cite{schroeder1991, koibuchi2001, cherkasova1996} do not use virtual channels to
achieve deadlock freedom, but turn prohibitions. 
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, then placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
While the hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{dally1994, duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, sui2000, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010, kim2006, park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present SR, an approach the solves these
limitations by allowing turn restrictions to be placed locally,
independently from other restrictions. The whole network is
partitioned into segments, and each bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom and connected networks. This \emph{locality
independence} property, together with no requirement of any particular tree/root
choice, would make it the best choice for the given scenario;
however, its topology independency still requires the knowledge of the
whole network graph in order to find the segments.

The DiSR approach presented in this work is a first attempt of
reaching the same goals of a segment based approach, that is
establishing segments, without requiring any topology graph or
centralized execution.
